---
title: 'Lab 1 Notes: Wrangling review and a few new things'
author: "German Silva"
date: "1/6/2022"
output: html_document
---

# Objectives

- Reminders of reproducible workflows
- Review some data wrangling basics (dplyr & tidyr)
- Introduce a couple new functions (separate, unite, drop_na, rename, relocate)
- Remember ggplot?
- A map in R

```{r setup, include=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# Attach packages
library(tidyverse)
library(here)
library(sf)
library(tmap)

### update.packages(ask = FALSE)
```

ctrl + shift + enter is shortcut to run whole code chunk

# Read in the data

```{r}
sf_trees <- read_csv(here("data", "sf_trees", "sf_trees.csv"))

```

# Part 1: Wrangling and ggplot review

**Examples 1:** find counts of observation by `legal_status` & wrangle a bit.

```{r}
# method 1: group_by() %>% summarize()

sf_trees %>% 
  group_by(legal_status) %>% 
  summarise(tree_count = n())

# method 2: different way plus a few new functions

top_5_status <- sf_trees %>% 
  count(legal_status) %>% # count by the column indicated here
  drop_na(legal_status) %>% # drop na's in this column
  rename(tree_count = n) %>% 
  relocate(tree_count) %>% 
  slice_max(tree_count, n = 5) %>% 
  arrange(-tree_count) # or arrange(desc(tree_count))

```

Review ggplot

```{r}
ggplot(top_5_status, aes(x= fct_reorder(legal_status, tree_count), y= tree_count))+
  geom_col(fill = 'darkgreen') +
  labs(x= 'Legal status',
       y= "Tree Count")+
  coord_flip() +
  theme_minimal()

```

**Example 2**: Only goin to keep observations where legal status is "Permitted Site" and caretaker is MTA and stored as `permitted_data_df`

```{r}
permitted_data_df <- sf_trees %>% 
  filter(legal_status == 'Permitted Site',
         caretaker == "MTA")
```

Example 3: Only keep Blackwood Acacia trees and then only keep columns legal_status, date, latitude, longitude, and store as `blackwood_acacia_df`

```{r}
blackwood_acacia_df <- sf_trees %>% 
  filter(str_detect(species, 'Blackwood Acacia')) %>% 
  select(legal_status, date, latitude, longitude)

##make a little graph of locations
ggplot(blackwood_acacia_df, aes(x=longitude, y= latitude))+
  geom_point(color = 'darkgreen')
```

**Example 4:** Use `tidyr::separate`

```{r}
sf_trees_sep <- sf_trees %>% 
  separate(species, into = c('spp_scientific', 'spp_common'), sep= ' :: ' )
```

Example 5: `tidyr::reunite`

```{r}
ex_5 <- sf_trees %>% 
  unite('id_status', tree_id, legal_status, sep = "_COOL_")
```

